---
title: "Upload Data to the Energy Databus with R"
author: "Anastasis Giannousakis"
date: "`r Sys.Date()`"
output: pdf_document
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Upload Data to the Energy Databus with R}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document describes an R tool for easy publishing of a dataset on
the Energy Databus (<https://energy.databus.dbpedia.org/>).

Extensive documentation of the Databus (including other methods for registering data)
can be found here:  
<https://github.com/LOD-GEOSS/databus-snippets/wiki>  
  
and here:  
<https://github.com/LOD-GEOSS/LOD-GEOSS.github.io/tree/master/documents>.

The R package `r2databus` contains the function `databusUpload` which will publish
your dataset on the Databus. The function awaits 2 arguments:
an API key (string; generated here:  
`https://energy.databus.dbpedia.org/{{user}}#settings`),
and an R list or a path to a JSON file with the dataset's metadata. An example of 
such metadata is shown here:
```
List of 14
 $ user             : chr "USERNAME"
 $ group            : chr "GROUPNAME"
 $ groupTitle       : chr "GROUPTITLE"
 $ groupAbstract    : chr "GROUPABSTRACT"
 $ groupDescription : chr "GROUPDESCRIPTION"
 $ artifact         : chr "ARTIFACT"
 $ version          : chr "YOURVERSION"
 $ dataidTitle      : chr "DATAIDTITLE"
 $ dataidAbstract   : chr "DATAIDABSTRACT"
 $ dataidDescription: chr "DATAIDDESCRIPTION"
 $ license          : chr "http://this.is/a-license"
 $ file             : chr "http://this.is/a-file-with-extension.ext"
 $ extension        : chr "EXT"
 $ metadata         : List of 2
  ..$ TYPE: chr "SOMEDATATYPE"
  ..$ FOO : chr "BAR"
```

All capitalized strings are mandatory (but you can keep the default values for 
testing), the same holds for the two paths `license`
and `file`. The `file` field has to be pointing to an existing file (URI). The 
`metadata` list can contain more than two items, and has to contain at least one.

## Example

```{r, eval = FALSE}
r2databus::databusUpload(myKey = "a-s3cr3t-K3y", myData = "/path/to/JSON/file")
```

## Further Reading

The Databus API essentially expects two JSON-LD files with very specific format,
which the `databusUpload` function will generate for you, and then post to the
Databus API. These are: a JSON-LD
file or string, containing the "group" information,
and another JSON-LD file or string, containing the "DataID".
They are explained in detail in the following.


### Group JSON-LD

JSON-LD files are JSON files (i.e. files in human- and machine-readble format)
containing **L**inked **D**ata. A 'group' on the databus can contain several DataID's.
It helps to group registered
data together. Here is an example of such file:

```
{
  "@context": "https://downloads.dbpedia.org/databus/context.jsonld",
  "@graph": {
    "@id": "https://energy.databus.dbpedia.org/username/GROUPNAME",
    "@type": "Group",
    "title": "Cleaned Data",
    "abstract": "Various cleaned datasets from DBpedia using AI.",
    "description": "Various cleaned datasets from DBpedia using AI."
  }
}
```



### DataID JSON-LD

A DataID contains one or more datasets and their description.
Here is an example of a DataID JSON-LD:

```
{
  "@context": "https://downloads.dbpedia.org/databus/context.jsonld",
  "@graph": {
  "@type": "Dataset",
  "@id": "https://energy.databus.dbpedia.org/giannoupik/GROUPNAME/ARTIFACT/YOURVERSION#Dataset",
  "hasVersion": "YOURVERSION",
  "title": "YOURTITLE",
  "abstract": "YOURABSTRACT",
  "description": "YOURDESCRIPTION",
  "license": "http://creativecommons.org/licenses/by/4.0/",
  "distribution": [
    {
      "@id": "https://energy.databus.dbpedia.org/user/group/artf/vers#artf_type=randomData.csv",
      "@type": "Part",
      "file": "https://energy.databus.dbpedia.org/user/group/artf/vers/artf_type=randomData.csv",
      "formatExtension": "csv",
      "compression": "none",
      "downloadURL": "https://yum-yab.github.io/data/databus-api-test/first/Sample500.csv",
      "byteSize": 34243,
      "sha256sum": "1c69ff99c105ab0f3459a4cd928f14284c996702148a2f62637df69f3e1a01ab",
      "dcv:type": "randomData"
    }
  ]
}
}
```
